Automated data-driven symbolic modeling of the lift forces acting on flapping wings

### 2021-01-18 paper/main.tex
Add subfigure captions
Reformat paper to use basic article class, and associated changes

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

@@

\title{Automated data-driven symbolic modeling of the lift forces acting on flapping wings}

\author[1,*]{Charles Richter}
\author[2,3]{Hod Lipson}
% \author[1,2,+]{Christine Author}
% \author[2,+]{Derek Author}
\affil[1]{Cornell University, Sibley School of Mechanical and Aerospace
Engineering, Ithaca, 14850, United States}
\affil[2]{Columbia University, Department of Mechanical Engineering, New York,
10027, United States}
\affil[3]{Columbia University, Data Science Institute, New York, 10027, United
States}

\affil[*]{car45@cornell.edu}

% \affil[+]{these authors contributed equally to this work}

@@

\begin{abstract}
We explore a data-driven approach to symbolically modeling the lift produced by
the flapping wing of a micro air vehicle.  The aerodynamics of flapping flight
pose a modeling challenge due to complex effects related to vortex dynamics,
wake interaction, aeroelasticity and three-dimensional flow, yet simple and
accurate equations are useful for engineering purposes and optimization of wing
design.  Experiments were conducted to measure the flapping kinematics and
instantaneous lift forces produced by eleven different 3D printed wing
geometries at three different flapping amplitudes and multiple speeds, totaling
68 distinct sets of data.  58 of these data sets were used to train data-driven
models using a symbolic regression algorithm, while the remaining sets were
reserved for model testing.  Data-driven models were found to be more accurate
and much simpler than existing quasi-steady analytical models consistent with
aerodynamic theory.  Furthermore, data-driven models were found to separate
distinct physical effects into inertial and aerodynamic components bearing
remarkable resemblance to their analytical counterparts. This result suggests
that a data-driven approach could be a powerful tool in separating and
identifying the structural building blocks of a wide variety of unknown
dynamical systems.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

@@

\subsection{Data-driven models}
The strategy for modeling lift forces symbolically was to train candidate equations on 58 of the 68 total experiments simultaneously, with the remaining ten reserved for testing.  These 58 training sets were intended to represent a diverse sampling of wing geometry, flapping speed and amplitude combinations.  The results from these 58 experiments were compiled into a single dataset listing roughly two periods of smoothed, registered flapping data per experiment.  The data consisted of instantaneous measured vertical force along with the measured value of each kinematic variable, the first and second derivatives of those variables, and the geometric parameters necessary to describe the wing's shape and distribution of area, totaling 31 separate variables.

@@

\subsection{Analytical models for comparison}
The analytical models used for comparison are also listed in Table~\ref{table:eqmodels}.  These models consist of calculated inertial forces added to a collection of aerodynamic terms.  The primary aerodynamic contribution in all quasi-steady models is a translational lift term proportional to the squared velocity of each span-element of the wing that is related to the wing's angle of attack and empirically determined coefficient of lift.  This term is labeled $L_{planar}$ when wing motions out of the nominal horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when out-of-plane motions are included. This translational lift term is given by:
\begin{equation}
L_{trans}=(\nicefrac{1}{2})\rho v^2AC_L,
\end{equation}
where \(v\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.  The theoretical value of the coefficient of lift depends on the angle of attack \(\alpha\):
\begin{equation}
C_L = C_{Lmax}\sin(2\alpha).
\end{equation}

@@

\subsection{Evaluation of Model Performance}
An evaluation of model performance allows comparison between analytical models and the equations developed through the data mining process.  Equations are judged by their accuracy and complexity.  In this case, the complexity or size of an equation was measured by simply counting the number of arithmetic and trigonometric operators used to combine the measured kinematic variables and geometric properties into an equation for instantaneous vertical force.  Both our data-driven models and the analytical equations are measured in this way.  For fairness, constants in the analytical models that would have been lumped together into larger coefficients by the symbolic regression algorithm (such as material density, air density, \(\pi\) or factors of integration) are also lumped together in the calculations of size for the analytical equations.  The size of the inertial calculations was added to the sizes of analytical lift calculations.  The 
pendulum motion has an equation size of 13 and the total inertial calculation including out-of-plane deviation has a size of 27.

@@

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_1}
\caption{\label{fig:eq_goodbad_1}}
\end{subfigure} &
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_2}
\caption{\label{fig:eq_goodbad_2}}
\end{subfigure} \\
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_3}
\caption{\label{fig:eq_goodbad_3}}
\end{subfigure} &
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_4}
\caption{\label{fig:eq_goodbad_4}}
\end{subfigure}
\end{tabular}
\caption{Comparison of force predictions between $EQ_4$ (blue) and
Equation~\eqref{eq:badeq} (red). $EQ_4$ was trained on 58 distinct experiments
simultaneously, including the data shown in~\subref{fig:eq_goodbad_1}, whereas
Equation~\eqref{eq:badeq} was trained \emph{only} on the data shown
in~\subref{fig:eq_goodbad_1}.  \subref{fig:eq_goodbad_1} shows that both models
perform well on this data within their training sets.  However, when presented
with test data \subref{fig:eq_goodbad_2}, \subref{fig:eq_goodbad_3} and
\subref{fig:eq_goodbad_4} not contained within either model's training set, only $EQ_4$ is able to
generalize and provide accurate predictions, while Equation~\eqref{eq:badeq}
fails to do so.}
\label{fig:eq_goodbad}
\end{figure}

\subsection{Performance of models trained on one experiment versus many experiments}
Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of two models on data within
their training sets, while Figures~\ref{fig:eq_goodbad_2},~\ref{fig:eq_goodbad_3},
and~\ref{fig:eq_goodbad_4} illustrate the performance of those same two models on data outside
their training sets. The differences between these two models result from the fact that one of them,
$EQ_4$ from Table~\ref{table:eqmodels} (drawn in blue), was trained on 58 separate experiments
simultaneously, whereas the other one, shown below (drawn in red), is one of many models trained
only on the single experiment shown in Figure~\ref{fig:eq_goodbad_1}:
\begin{align}
 Force = \frac{C_1 + C_2\psi^2 - C_3\psi}{C_4 + \ddot{\theta} + C_5\dot{\theta} + \cos(C_6 -
  \frac{C_7}{\psi^2} - C_8\psi)} - C_9.
\label{eq:badeq}
\end{align}
Thus, $EQ_4$ shows strong qualitative and quantitative agreement with all of the experiments in Figure~\ref{fig:eq_goodbad}, indicating that this model is not over-fit but in fact generalizes well to new data. Equation~\eqref{eq:badeq}, on the other hand, fails to accurately predict the recorded data outside its training set.  Insufficient data and over-fitting are potential pitfalls of data-driven modeling, but are overcome in this case with a diverse sampling of training data.

@@

\subsection{Correspondence of physical terms}
One powerful result of the data-driven modeling approach is that in many cases, certain terms from the data-driven models closely match physically understandable terms from the analytical models.  The correspondence between these terms is listed in Table~\ref{table:eqcorrespondence}. Taking model $EQ_4$ as an example, all three terms show strong resemblance to inertial and aerodynamic components of the analytical models.  Figure~\ref{fig:eqcomparison} shows a comparison between the simple velocity-squared aerodynamic lift prediction and the second term of $EQ_4$: \(C_2RX_{CM}\psi\dot{\phi}\).  These two calculations are numerically very similar, yet the third term of $EQ_4$ is much simpler than $L_{trans}$.

@@

\begin{figure}[ht]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_comparison_lift}
\caption{\label{fig:eqcomparison_lift}}
\end{subfigure}
% \qquad
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_comparison_inertial}
\caption{\label{fig:eqcomparison_inertial}}
\end{subfigure}
\caption{Comparison between terms of the analytical expressions (red) and data-driven models (blue).
  \subref{fig:eqcomparison_lift} Primary translational aerodynamic lift proportional to the square
  of the wing velocity. \subref{fig:eqcomparison_inertial} Inertial forces related to the pendulum
  motion of the wing and deviation out of the stroke plane. The modeling process automatically
  decoupled aerodynamic and inertial effects with no prior knowledge of the system. The total
  measured force is shown for reference (black).}
\label{fig:eqcomparison}
\end{figure}

@@

\subsection{3D Printed wings and actuation mechanism}
The experimental wings were fabricated using an Objet Connex500 3D printer with
the rigid Objet FullCure720 material.
Flapping flight with 3D printed wings was reported in \cite{richter2011untethered}. The
wing design consisted of an interlocking root and tip, allowing the wing
surface to pivot naturally during flapping from its neutral vertical position
to an angle of attack of approximately 45\textdegree. The wing design is illustrated in
Figure~\ref{fig:wing1} and the hinge design is
illustrated in Figure~\ref{fig:root2}.  In total, eleven wings were used,
ranging in wingspan from 40 mm to 120 mm and ranging in chord length from 40 mm
to 80 mm. Flapping amplitudes were constrained by the geometry of the wing root
design to be set at 72\textdegree, 81\textdegree, and 96\textdegree, and each
wing was flapped at speeds ranging between 1 Hz and 3.5 Hz for each of these
amplitude settings.  68 sets of data were recorded, representing Reynolds
numbers between 2500 and 5000.

@@

\subsection{Measurement apparatus}
The apparatus used to measure the forces and kinematics of flapping wings was
based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
logging.  This force measurement system was capable of a sampling rate of 1 kHz
and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
a crank and connecting rod used to drive the experimental wings in a roughly
sinusoidal flapping motion. This system is pictured in
Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency
of the load cell and wing driving mechanism was roughly 70 Hz, the natural
frequency for the first bending mode of the experimental wing was approximately
26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
All experimental force measurements were filtered using a high-order zero-phase
digital low-pass filter with a cutoff frequency of 7.5 times the flapping
frequency, which preserved the primary inertial and aerodynamic forces acting
on the wings while eliminating the faster vibrations of the wing structure and
the load cell apparatus.

@@

\subsection{High-speed measurement of flapping kinematics}
A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
210 frames per second from a top-down perspective as well as a side view
provided by a mirror mounted in the field of view.  The main beam of each wing
was colored red to enable precise tracking of the stroke angle.  The top-down
projected area of the wing was used to extract deflection angle, and tracking
of the wing tip in the side-view was used to extract the angle of deviation out
of the wing's stroke plane.  These three angles precisely define the
configuration of the wing.  A schematic diagram of this apparatus is shown in
Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
to extract the kinematics of the wing motion from high-speed video.  The
red-colored main beam was identified with a color filter and a line was fitted
using a Hough Transform to identify the stroke angle.  The wing deflection
angle was computed by measuring the projected area of the wing in the image and
compared with the known dimensions of the rectangular wing shape. The
out-of-plane deviation angle of the wing was measured by tracking the leading
and trailing edge corners of the wingtip in the side-view of the mirror. We refer the reader to our
previous work for a more detailed description of our methods of extracting wing kinematics from
high speed video and processing the measured data\cite{richter2012thesis}.
% Figure~\ref{fig:kinematics} provides sample force data and measured wing kinematics from one
% experiment.

@@

\bibliography{main}

>>>>

\documentclass{article}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{authblk}
\usepackage{fullpage}

% \documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

@@

% Keywords command
\providecommand{\keywords}[1]
{
  \vspace{5mm}
  \small
  \noindent
  \textbf{Keywords:} #1
  \normalsize
}

\title{Automated data-driven symbolic modeling of the lift forces acting on flapping wings}
\date{}

\author[1,*]{Charles Richter}
\author[2,3]{Hod Lipson}
\affil[1]{Cornell University, School of Mechanical and Aerospace Engineering, Ithaca, United States}
\affil[2]{Columbia University, Department of Mechanical Engineering, New York, United States}
\affil[3]{Columbia University, Data Science Institute, New York, United States}
\affil[*]{car45@cornell.edu}

\begin{document}
\maketitle

\begin{abstract}
We explore a data-driven approach to symbolically modeling the lift produced by
the flapping wing of a micro air vehicle.  The aerodynamics of flapping flight
pose a modeling challenge due to complex effects related to vortex dynamics,
wake interaction, aeroelasticity and three-dimensional flow, yet simple and
accurate equations are useful for engineering purposes and optimization of wing
design.  Experiments were conducted to measure the flapping kinematics and
instantaneous lift forces produced by eleven different 3D printed wing
geometries at three different flapping amplitudes and multiple speeds, totaling
68 distinct sets of data.  58 of these data sets were used to train data-driven
models using a symbolic regression algorithm, while the remaining sets were
reserved for model testing.  Data-driven models were found to be more accurate
and much simpler than existing quasi-steady analytical models consistent with
aerodynamic theory.  Furthermore, data-driven models were found to separate
distinct physical effects into inertial and aerodynamic components bearing
remarkable resemblance to their analytical counterparts. This result suggests
that a data-driven approach could be a powerful tool in separating and
identifying the structural building blocks of a wide variety of unknown
dynamical systems.
\end{abstract}

@@

% \flushbottom
% \maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
% \thispagestyle{empty}

@@

\subsection*{Data-driven models}
The strategy for modeling lift forces symbolically was to train candidate equations on 58 of the 68 total experiments simultaneously, with the remaining ten reserved for testing.  These 58 training sets were intended to represent a diverse sampling of wing geometry, flapping speed and amplitude combinations.  The results from these 58 experiments were compiled into a single dataset listing roughly two periods of smoothed, registered flapping data per experiment.  The data consisted of instantaneous measured vertical force along with the measured value of each kinematic variable, the first and second derivatives of those variables, and the geometric parameters necessary to describe the wing's shape and distribution of area, totaling 31 separate variables.

@@

\subsection*{Analytical models for comparison}
The analytical models used for comparison are also listed in Table~\ref{table:eqmodels}.  These models consist of calculated inertial forces added to a collection of aerodynamic terms.  The primary aerodynamic contribution in all quasi-steady models is a translational lift term proportional to the squared velocity of each span-element of the wing that is related to the wing's angle of attack and empirically determined coefficient of lift.  This term is labeled $L_{planar}$ when wing motions out of the nominal horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when out-of-plane motions are included. This translational lift term is given by:
\begin{equation}
L_{trans}=(\nicefrac{1}{2})\rho v^2AC_L,
\end{equation}
where \(v\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.  The theoretical value of the coefficient of lift depends on the angle of attack \(\alpha\):
\begin{equation}
C_L = C_{Lmax}\sin(2\alpha).
\end{equation}

@@

\subsection*{Evaluation of Model Performance}
An evaluation of model performance allows comparison between analytical models and the equations developed through the data mining process.  Equations are judged by their accuracy and complexity.  In this case, the complexity or size of an equation was measured by simply counting the number of arithmetic and trigonometric operators used to combine the measured kinematic variables and geometric properties into an equation for instantaneous vertical force.  Both our data-driven models and the analytical equations are measured in this way.  For fairness, constants in the analytical models that would have been lumped together into larger coefficients by the symbolic regression algorithm (such as material density, air density, \(\pi\) or factors of integration) are also lumped together in the calculations of size for the analytical equations.  The size of the inertial calculations was added to the sizes of analytical lift calculations.  The 
pendulum motion has an equation size of 13 and the total inertial calculation including out-of-plane deviation has a size of 27.

@@

\begin{figure}[ht]
\centering
\begin{tabular}{cc}
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_1}
\caption{\label{fig:eq_goodbad_1} Training data.}
\end{subfigure} &
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_2}
\caption{\label{fig:eq_goodbad_2} Test data \#1.}
\end{subfigure} \\
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_3}
\caption{\label{fig:eq_goodbad_3} Test data \#2.}
\end{subfigure} &
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_4}
\caption{\label{fig:eq_goodbad_4} Test data \#3.}
\end{subfigure}
\end{tabular}
\caption{Comparison of force predictions between $EQ_4$ (blue) and
Equation~\eqref{eq:badeq} (red). $EQ_4$ was trained on 58 distinct experiments
simultaneously, including the data shown in~\subref{fig:eq_goodbad_1}, whereas
Equation~\eqref{eq:badeq} was trained \emph{only} on the data shown
in~\subref{fig:eq_goodbad_1}.  \subref{fig:eq_goodbad_1} shows that both models
perform well on this data within their training sets.  However, when presented
with test data \subref{fig:eq_goodbad_2}, \subref{fig:eq_goodbad_3} and
\subref{fig:eq_goodbad_4} not contained within either model's training set, only $EQ_4$ is able to
generalize and provide accurate predictions, while Equation~\eqref{eq:badeq}
fails to do so.}
\label{fig:eq_goodbad}
\end{figure}

\subsection*{Performance of models trained on one experiment versus many experiments}
Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of two models on data within
their training sets, while Figures~\ref{fig:eq_goodbad_2},~\ref{fig:eq_goodbad_3},
and~\ref{fig:eq_goodbad_4} illustrate the performance of those same two models on data outside
their training sets. The differences between these two models result from the fact that one of them,
$EQ_4$ from Table~\ref{table:eqmodels} (drawn in blue), was trained on 58 separate experiments
simultaneously, whereas the other one, shown below (drawn in red), is one of many models trained
only on the single experiment shown in Figure~\ref{fig:eq_goodbad_1}:
\begin{align}
 Force = \frac{C_1 + C_2\psi^2 - C_3\psi}{C_4 + \ddot{\theta} + C_5\dot{\theta} + \cos(C_6 -
  \frac{C_7}{\psi^2} - C_8\psi)} - C_9.
\label{eq:badeq}
\end{align}
Thus, $EQ_4$ shows strong qualitative and quantitative agreement with all of the experiments in Figure~\ref{fig:eq_goodbad}, indicating that this model is not over-fit but in fact generalizes well to new data. Equation~\eqref{eq:badeq}, on the other hand, fails to accurately predict the recorded data outside its training set.  Insufficient data and over-fitting are potential pitfalls of data-driven modeling, but are overcome in this case with a diverse sampling of training data.

@@

\subsection*{Correspondence of physical terms}
One powerful result of the data-driven modeling approach is that in many cases, certain terms from the data-driven models closely match physically understandable terms from the analytical models.  The correspondence between these terms is listed in Table~\ref{table:eqcorrespondence}. Taking model $EQ_4$ as an example, all three terms show strong resemblance to inertial and aerodynamic components of the analytical models.  Figure~\ref{fig:eqcomparison} shows a comparison between the simple velocity-squared aerodynamic lift prediction and the second term of $EQ_4$: \(C_2RX_{CM}\psi\dot{\phi}\).  These two calculations are numerically very similar, yet the third term of $EQ_4$ is much simpler than $L_{trans}$.

@@

\begin{figure}[ht]
\centering
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_comparison_lift}
\caption{\label{fig:eqcomparison_lift} Aerodynamic lift forces.}
\end{subfigure}
% \qquad
\begin{subfigure}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/eq_comparison_inertial}
\caption{\label{fig:eqcomparison_inertial} Inertial forces.}
\end{subfigure}
\caption{Comparison between terms of the analytical expressions (red) and data-driven models (blue).
  \subref{fig:eqcomparison_lift} Primary translational aerodynamic lift proportional to the square
  of the wing velocity. \subref{fig:eqcomparison_inertial} Inertial forces related to the pendulum
  motion of the wing and deviation out of the stroke plane. The modeling process automatically
  decoupled aerodynamic and inertial effects with no prior knowledge of the system. The total
  measured force is shown for reference (black).}
\label{fig:eqcomparison}
\end{figure}

@@

\subsection*{3D Printed wings and actuation mechanism}
The experimental wings were fabricated using an Objet Connex500 3D printer with
the rigid Objet FullCure720 material.
Flapping flight with 3D printed wings was reported in \cite{richter2011untethered}. The
wing design consisted of an interlocking root and tip, allowing the wing
surface to pivot naturally during flapping from its neutral vertical position
to an angle of attack of approximately 45\textdegree. The wing design is illustrated in
Figure~\ref{fig:wing1} and the hinge design is
illustrated in Figure~\ref{fig:root2}.  In total, eleven wings were used,
ranging in wingspan from 40 mm to 120 mm and ranging in chord length from 40 mm
to 80 mm. Flapping amplitudes were constrained by the geometry of the wing root
design to be set at 72\textdegree, 81\textdegree, and 96\textdegree, and each
wing was flapped at speeds ranging between 1 Hz and 3.5 Hz for each of these
amplitude settings.  68 sets of data were recorded, representing Reynolds
numbers between 2500 and 5000.

@@

\subsection*{Measurement apparatus}
The apparatus used to measure the forces and kinematics of flapping wings was
based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
logging.  This force measurement system was capable of a sampling rate of 1 kHz
and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
a crank and connecting rod used to drive the experimental wings in a roughly
sinusoidal flapping motion. This system is pictured in
Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency
of the load cell and wing driving mechanism was roughly 70 Hz, the natural
frequency for the first bending mode of the experimental wing was approximately
26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
All experimental force measurements were filtered using a high-order zero-phase
digital low-pass filter with a cutoff frequency of 7.5 times the flapping
frequency, which preserved the primary inertial and aerodynamic forces acting
on the wings while eliminating the faster vibrations of the wing structure and
the load cell apparatus.

@@

\subsection*{High-speed measurement of flapping kinematics}
A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
210 frames per second from a top-down perspective as well as a side view
provided by a mirror mounted in the field of view.  The main beam of each wing
was colored red to enable precise tracking of the stroke angle.  The top-down
projected area of the wing was used to extract deflection angle, and tracking
of the wing tip in the side-view was used to extract the angle of deviation out
of the wing's stroke plane.  These three angles precisely define the
configuration of the wing.  A schematic diagram of this apparatus is shown in
Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
to extract the kinematics of the wing motion from high-speed video.  The
red-colored main beam was identified with a color filter and a line was fitted
using a Hough Transform to identify the stroke angle.  The wing deflection
angle was computed by measuring the projected area of the wing in the image and
compared with the known dimensions of the rectangular wing shape. The
out-of-plane deviation angle of the wing was measured by tracking the leading
and trailing edge corners of the wingtip in the side-view of the mirror. We refer the reader to our
previous work for a more detailed description of our methods of extracting wing kinematics from
high speed video and processing the measured data\cite{richter2012thesis}.
% Figure~\ref{fig:kinematics} provides sample force data and measured wing kinematics from one
% experiment.

@@

\bibliographystyle{abbrv}
\bibliography{main}
